---
title: "Assignment 2"
output: html_document
---

Assignment 2: Weather Problems
========================================================
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.
This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

Installing and loading the required packages:
```{r, echo = FALSE}
install.packages("plyr")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("knitr")

library("plyr")
library("dplyr")
library("ggplot2")
library("knitr")
```

Getting the data:
```{r}
#download.file("https://d396qusza40orc.cloudfront.net/repdata/data/StormData.csv.bz2", "StormData.csv.bz2")
stormData <- read.csv("StormData.csv.bz2")
head(stormData)
```

# Data Processing
## Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?
First, in order to calculate the most harmful events, the raw data from the stormData dataset is summarized by its two main observations: fatalities and injuries.
```{r}
casualties <- ddply(stormData, .(EVTYPE), summarize, fatalities = sum(FATALITIES), injuries = sum(INJURIES))
```
Then, both observations were added and the subset was ordered desdending.
```{r}
casualties$totalCasualties <- casualties$fatalities + casualties$injuries
casualties <- casualties[with(casualties, order( - totalCasualties)),][1:5,]
ggplot(casualties, aes(x = EVTYPE, y = totalCasualties)) + geom_bar(stat = "identity", fill = "light blue") + xlab("Event type") + ylab("Total casualties")
```

##Across the United States, which types of events have the greatest economic consequences?
First, in order to calculate the events with greatest economic consequences, the raw data from the stormData dataset is summarized by its two main observations: property and crop damages.
```{r}
damages <- ddply(stormData, .(EVTYPE), summarize, property = sum(PROPDMG), crop = sum(CROPDMG))
```
Then, both observations were added and the subset was ordered desdending.
```{r}
damages$totalDamages <- damages$property + damages$crop
damages <- damages[with(damages, order( - totalDamages)),][1:5,]
ggplot(damages, aes(x = EVTYPE, y = totalDamages)) + geom_bar(stat = "identity", fill = "light blue") + xlab("Event type") + ylab("Total damages ($)")
```

# Results 
- Across the United States, TORNADOES are the most harmful events.
- Across the United States, FLOODS are the events with the greatest economic consequences.